On the minimum leaf conductance: its role in models of plant water use, and ecological and environmental controls

### 2018-01-10 manuscript.Rmd
much writing

```{r, echo=FALSE}
tab <- read.csv(text="I. Introduction
II. Comparison of various definitions of minimum conductance
III. Problems with estimating minimum conductance from cuvette data
IV. Use of minimum conductance in models
\t1. Models of water-use efficiency
\t2. Models of drought response of stomatal conductance
V. Cuticular conductance
VI. Nighttime conductance
VII. Environmental and ecological variation in minimum conductance 
\t1. Contribution of stomata to minimum conductance
\t2. Minimum conductance shows little meaningful interspecific variation
\t3. Acclimation to drought
\t4. Acclimation and instantaneous response to temperature
\t5. Other factors
VIII. Conclusions", header=F)

@@

\
Stomata exert a strong control on the transpiration rate of leaves [@bange1953]. It is well known that stomata respond quickly to changes in light intensity, humidity and CO~2~ concentration at the leaf surface. In doing so, stomata can be thought to balance the uptake of CO~2~ with the loss of water, and thus can be expected to close stomata when photosynthetic rates are low. To avoid desiccation and ultimate death, stomata can also be expected to close during periods of severe water stress. Plants cannot, however, completely halt water loss even when stomata are closed.

\
The absolute minimum attainable water loss rate through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf. Other measurements allow for the estimation of water loss through incompletely closed and broken stomata. Of particular interest is the method where leaves are detached from the plant, and their weight loss monitored over time. This method (mass loss of detached leaves, MLD) is of particular interest because it more closely simulates actual field conditions for plants during severe drought, and includes both surfaces of the leaf - not just the stomata-free side which is typically used for cuticular conductance measurements.

\
Most current-generation process-based global vegetation models (GVMs) use a version of the Ball-Berry class of stomatal conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity.

@@

\
Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), and g~0~ is the value of g~s~ when A~n~ is zero. In this paper, we refer to Eq. 1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987). Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been largely ignored. For example, a recent comprehensive review of stomatal conductance models did not mention g~0~ or any similar minimum conductance [@damour2010]. The original description of Eq. 1 as published by @ballberry1987 did not include an intercept term. It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model. In practice, g~0~ is estimated from a regression of Eq. 1 with leaf gas exchange data, but this may lead to inaccurate values, as pointed out by @barnard2013 but not explained in detail.

\
We thus have several different definitions of what constitutes a 'minimum' conductance of leaves. In this review, we demonstrate that inclusion of a minimum conductance is necessary both for simulation of the response of water-use efficiency to various environmental drivers, as well as a component of the response of plants to severe drought. However, there is no guidance at the moment on how to estimate minimum conductance for use in models, how various definitions compare quantitatively, and whether it is feasible to assign species-specific values of the minimum conductance. We do not expect that g~0~, cuticular conductance (g~cut~) and minimum conductance from mass loss of detached leaves (we refer to minimum conductance measured with this method further as g~min~) to give similar values, in fact the contribution of stomata to each of these conductances will be very different [@brodribb2014].

\
Previews reviews of the minimum conductance have largely focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles [@riederer2008, @schuster2017; @kerstiens1996; @fernandez2017]. Other work has focused on the role of g~0~ in models of plant water use [e.g. @barnard2013; @lombardozzi2017], without referring to the literature on cuticular conductance. Thus there is a disconnect between the dizzying level of understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models. We aim to improve this connection by synthesizing the state of knowledge, thereby pointing out gaps in understanding, and suggest areas for further research.

@@

\
Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories (see Table 1). The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cut~) are typically made on isolated non-stomatal cuticles [@riederer2001], via gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata (g~cut,seal~) [@kerstiens1996]. The method we here refer to as minimum conductance is that of mass loss of intact detached leaves (MLD). In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out. After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit [@sinclair1986]. This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress. Finally, the g~0~ parameter, which is a model-based minimum conductance reached when photosynthesis rate is zero, is typically estimated by applying regression to Eq. 1 (of which various forms exist).

\ 
We quantitatively compared each of these different 'minimum' conductance values by synthesizing existing data. To estimate g~0~ from regression, we used the @lin2015 database to fit the model of @medlyn2011 using non-linear regression, and data presented by @miner2017 on the linear regression of the @ballberry1987 model. Measurements of conductance of isolated, non-stomatal cuticles (g~cut,isol~), and of conductance of detached leaves with the stomatal side sealed (g~cut,seal~) were taken from the compilation by @kerstiens1996.  Data on minimum conductance (g~min~) from MLD was compiled from 44+ original papers (see SuppInfo A for a list of papers and additional figures describing this new compilation). For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources. All conductance values were converted to per unit projected surface area. Units were converted to molar fluxes assuming 41 mol m^-3^ (cf. @kerstiens1996). Although it is quite clear that night-time conductance (g~dark~) is much larger than a true 'minimum' conductance, it is frequently used as an independent measurement of g~0~ [@lombardozzi2017], and thus a comparison of magnitudes is useful. A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca. 150 species during night-time. Actual minimum daytime values were estimated from the Lin et al. database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset. This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted/extrapolated g~0~ term.

\
(*expand*) The results are presented in Fig. 1. Conductance in the dark was on average eight-fold higher than g~min~. It is no new observation that stomata do not close in the dark, but do close on detached leaves, as already demonstrated by @hygen1954 in Scots pine, and described by @koerner1994. More recently, @walden-coleman2013 showed much higher g~dark~ compared to g~min~ in cotton.

However, @cavender-bares2007 gdark similar to gmin in oaks.

Actual measured stomatal conductance during low photosynthesis (g~s,low A~) was also much lower than fitted g~0~ or g~dark~. In other words, actual daytime conductance is much lower than the presumed minimum estimated from fitted g~0~. A similar result was already presented by @dekauwe2015 (their Fig. A2).

@@

**Figure 1**. Comparison of various estimates of the minimum conductance. Error bars are 95\% confidence intervals. Different letters denote significant differences (at $\alpha$ = 0.05). See Table 1 for definitions of minimum conductance, and text for the data sources.

@@

\
The standard approach in estimating g~0~ for use in Eq. 1 is to use regression with g~s~ as the response variable, and the right-hand side of Eq. 1 as the predictor. The g~0~ parameter is thus estimated as the intercept. Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated by this method. The difficulty of the estimation can already be anticipated from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g. @leuning1995 uses 0.01 mol m^-2^ s^-1^ for all species); presumably because unreliable estimates were obtained, b) negative g~0~ estimates from regression are commonly reported [@leuning1995; @heroult2013; @miner2017], though it is clearly non-sensical to suggest negative conductance values. @barnard2013 also mentioned the difficulty of fitting Eq. 1 to estimate g~0~, but did not present specific details.

@@

\
In Fig. 2, we demonstrate how g~0~ and g~1~ are statistically correlated, when fitted for a single study. The correlation between slopes and intercepts is not unique to Eq. 1, but a general property of linear models [@stapleton1995]. The consequence is that it is more difficult to interpret g~1~ and g~0~ independently, and that confidence intervals on each of the parameters should also not be interpreted independently (see @becker2007 for more discussion on the statistical aspects of comparing slopes and intercepts). (*expand*)

@@

**Figure 3**. Correlation between estimated g~0~ and the R^2^ of the regression. In the left panel, we fitted the linearized form of the Medlyn et al. (2011) model to each of the datasets in the Lin et al. (2015) database. Vertical lines are 95% confidence intervals. In the right panel, a very similar association between g~0~ and R^2^ is presented by the review of @miner2017, using the Ball-Berry model. In panel (c) coefficient of variation and R^2^.

@@

\
In this review we focus on the cuticle as a barrier to water loss, but point out that previous work has shown that the cuticle is much less permeable to CO~2~ compared to H~2~O [@boyer1997; @boyer2015]. This finding, if indeed generally true, has a large effect on the calculation of various gas exchange parameters [@hanson2006]. @manzoni2011 recalculated leaf water-use efficiency for droughted plants assuming CO~2~ is blocked by the cuticle, but the uncertainty of this approach is leaky and incompletely closed stomata (see Section XXX). Because g~0~ includes not just the cuticle, but also incompletely closed stomata, and the finding that g~0~ is much larger than estimates of g~min~ or g~cut~ (Fig. 1), we ignore this effect in the model simulations that follow.

\
Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero. But A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, begging the question whether the same g~0~ is reached in each of these cases. In the application of the Ball-Berry models, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005]. It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cut~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata (see Section XXX).

\
Photosynthesis also ceases at very high temperature (ref), but stomata do not appear to close in proportion to the decrease in photosynthesis, if at all [@drake, @urban2017], thus we cannot use high temperature to estimate a reasonable g~0~. Simulations of water use during heat waves depends highly on the value of g~0~ used [@kala2016] . It should be pointed out that the response of g~s~ to T~leaf~ is likely poorly modelled with a Ball-Berry model, since a number of studies have shown that g~s~ and A~n~ decouple at high T~leaf~, resulting in high rates of g~s~ when A~n~ is near zero [@urban2017] (Drake et al., unpublished).

@@

\
Using the coupled leaf gas exchange model, we show in Fig. 4 how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^). We also demonstrate the effect of g~0~ on modelling the T~leaf~ response. In this simulation, T~leaf~ and D are assumed to co-vary with an empirical relationship as used by @duursma2014. Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increase leaf transpiration E~L~, but at a slightly higher rate than just due to g~0~. This effect arises because the additional conductance allows slightly higher rates of photosynthesis, which in turn increase g~s~ via Eq. 1.

@@

**Figure 4**. Simulations  of intrinsic water-use efficiency (A~n~/g~s~), intercellular CO~2~ concentration (C~i~) and leaf transpiration rate (E~L~) with three values of g~0~. For the top two panels, all environmental drivers were held constant except the photosynthetic photon flux density (PPFD), demonstrating that the inclusion of g~0~ is necessary to simulate the decrease in A~n~/g~s~ at low PPFD. For the bottom two panels, VPD and T~air~ were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions. For all simulations it is assumed that T~leaf~ is equal to T~air~.

## 2. Models of drought response of stomatal conductance

\
Another reason to include a non-zero g~0~ term in models of stomatal conductance arises during severe drought. When plants are sufficiently water-stressed that stomata are completely closed, water loss still continues through the cuticle and leaky and incompletely closed stomata. Thus models that aim to predict when plants desiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000]. This is particularly important given the limited variation in stomatal response to drought among plants @martinstpaul2017.

@pisekwinkler1953, early calculation of the length of time needed to desiccate leaves to some water content, given the minimum transpiration rate, and the water content of the leaves.

@sinclair2000 gmin as a drought tolerance trait, can calculate time needed to death during drought phase when plants no longer take up water.

More recently, @gleason2014 and @blackman2016 proposed that embolism resistance along with whole plant capacitance and minimum transpiration rates all contribute to define the time to desiccation.

@burghardt2006 reported on a correlation between g~min~ and the survival time of leaves.

Finally, Martin-StPaul et al 2017 demonstrated in a whole plant model of hydraulic failure that gmin was probably one of the key parameters to explain water potential drops below cavitation thresholds during drought given the relatively conserved range of stomatal closure to drought.  Relying on the Sureau model presented in @martinstpaul2017 we illustrate the critical role of gmin in defining desiccation time.

@@

**Figure 5**

@@

\
Literature values of g~cut~, especially when measured on isolated cuticles (g~cut~) were much lower than any other estimate of conductance (Fig. 1), in line with the methods comparison of @kerstiens1996. More recently, @schuster2017 argued, based on a data for 8 species, that g~min~ from MLD is comparable to g~cut~. However in the same study, a literature compilation of hundreds of values showed that g~cut~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cut~ were pooled in their analysis. Nonetheless, the comparison between g~cut~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions. @boyer1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases.

\
A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, as these processes have been well described elsewhere [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b]. However, a few key points should be reiterated as they are relevant to the current discussion, in particular when we aim to interpret variation in literature values of g~min~.

\
Although we use the term cuticular conductance freely, this transport pathway does not represent a true conductance since water does not diffuse as a gas through the cuticle. Instead, it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle [@kerstiens1996; @riederer2001; @schreiber1996]. The main barrier to diffusion is actually a very thin layer at the leaf surface. Because most of the resistance is located in such a thin layer, g~cut~ does not correlate with the thickness of the cuticle  [@priestley1943; @riederer2001; @anfodillo2002; @schuster2016b]. Nonetheless there is considerable variation in cuticle thickness between plant species [@schuster2016a], as well as within species in response to drought (cite), along altitudinal transects [@delucia1984], and even with increasing height in the canopy of very tall trees [@woodruff2010]. If the cuticle thickness does not directly affect its conductance, what is the value of increasing the cuticle?

\
It is important to bear in mind the many other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, various interactions with insects including signalling and herbivory resistance [@mueller2008; @kerstiens1996c;@riederer2008]. The thickness of the cuticle may also confer mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness). Thick cuticles may also be useful because exposure to wind has been shown to dislodge cuticular compounds by abrasion [@rogge1993], thereby increasing the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991]. Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986], most dramatically demonstrated in an ice storm, leading to much higher g~min~ [@boyce2003]. Such exposure to the elements was also held responsible for thicker cuticles in coastal, wind-exposed sites compared to inland sites [@blackman2005].

>>>>

```{r, echo=FALSE}
tab <- read.csv(text="I. Introduction
II. Comparison of various definitions of minimum conductance
III. Problems with estimating minimum conductance from cuvette data
IV. Use of minimum conductance in models
\t1. Models of water-use efficiency
\t2. Models of plant desiccation
V. Cuticular conductance
VI. Nighttime conductance
VII. Environmental and ecological variation in minimum conductance 
\t1. Contribution of stomata to minimum conductance
\t2. Minimum conductance shows little meaningful interspecific variation
\t3. Acclimation to drought
\t4. Acclimation and instantaneous response to temperature
\t5. Other factors
VIII. Conclusions", header=F)

@@

\
Stomata exert a strong control on the transpiration rate of leaves [@bange1953]. It is well known that stomata respond quickly to changes in light intensity, humidity and CO~2~ concentration at the leaf surface. In doing so, stomata can be thought to balance the uptake of CO~2~ with the loss of water, and thus can be expected to close stomata when photosynthetic rates are low. To avoid desiccation and ultimate death, stomata can also be expected to close during periods of severe water stress. Plants cannot, however, completely halt water loss even when stomata are closed. There is increasing recognition that the minimum conductance plays an important role in estimating water fluxes in plant canopies [@barnard2013], during heat waves [@kala2016], and in models of plant drought response [@blackman2016; @martinstpaul2017].

\
Most current-generation process-based global vegetation models (GVMs) use a version of the Ball-Berry class of stomatal conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity (Eq. 1).

@@

\
Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), and g~0~ is the value of g~s~ when A~n~ is zero. In this paper, we refer to Eq. 1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987). Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been studied in much less detail. For example, a recent comprehensive review of stomatal conductance models did not mention g~0~ or any similar minimum conductance [@damour2010]. The original description of Eq. 1 as published by @ballberry1987 did not include an intercept term. It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model. In practice, g~0~ is estimated from a regression of Eq. 1 with leaf gas exchange data, but this may lead to inaccurate values, as pointed out by @barnard2013 but not explained in detail.

\
In this review, we demonstrate that inclusion of a minimum conductance is necessary both for simulation of the response of water-use efficiency to various environmental drivers, as well as a component of the response of plants to severe drought. However, there is no guidance at the moment on how to estimate minimum conductance for use in models, how various definitions and measurements compare quantitatively, whether it is feasible to assign species-specific values of the minimum conductance, and what are the ecological and environmental controls on g~min~ [@schuster2017].

\
The absolute minimum attainable water loss rate is through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf. Other measurements allow for the estimation of water loss through incompletely closed and broken stomata. Of particular interest is the method where leaves are detached from the plant, and their weight loss monitored over time. This method (mass loss of detached leaves, MLD) is interesting because it more closely simulates actual field conditions for plants during severe drought, and includes both surfaces of the leaf - not just the stomata-free side which is typically used for cuticular conductance measurements. A comprehensive compilation of estimates of g~min~ is currently lacking, and is needed not just to parameterize models, but also to study sources of variation in this overlooked plant trait.

\
Previous reviews of the minimum conductance have largely focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles [@riederer2008, @schuster2017; @kerstiens1996; @fernandez2017]. Other work has focused on the role of g~0~ in models of plant water use [e.g. @barnard2013; @lombardozzi2017], without referring to the literature on cuticular conductance. Thus there is a disconnect between the dizzying level of understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models. We aim to improve this connection by synthesizing the state of knowledge, thereby pointing out gaps in understanding, and suggest areas for further research. Our goals are to a) compare various definitions of g~min~ quantitively, which have been previously assumed equal, b) demonstrate the need to include a non-zero g~0~ (Eq. 1), for models of water-use efficiency as well as the drought response, c) present a new compilation of minimum conductance measurements, and demonstrate that the large acclimatory potential of this trait to environmental drivers.

@@

\
Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories (see Table 1). The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cut~) are typically made on isolated non-stomatal cuticles [@riederer2001], via special gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata [@kerstiens1996]. We further discuss cuticular conductance in Section V.

\
The minimum conductance of intact leaves is typically measured by mass loss of detached leaves (MLD). In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out. Early work by @hygen1951 showed that after a leaf is detached, initial water loss rates are high, but after a while a constant low rate is achieved. From this minimum transpiration rate (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), the minimum conductance (g~min~) can be estimated using measured vapour pressure deficit [@sinclair1986]. This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress.

\ 
We quantitatively compared each of the different 'minimum' conductance (Table 1) values by synthesizing existing data. To estimate g~0~ (Eq. 1) from regression, we used the @lin2015 database to fit the model of @medlyn2011 using non-linear regression, and data presented by @miner2017 on the linear regression of the @ballberry1987 model. Measurements of conductance of isolated, non-stomatal cuticles (g~cut~) were taken from the compilations by @kerstiens1996 and @schuster2017.  Data on minimum conductance (g~min~) from MLD was compiled from 44+ original papers (see SuppInfo 2 for a list of papers). For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources. All conductance values were converted to per unit projected surface area. Units were converted to molar fluxes assuming 41 mol m^-3^ (cf. @kerstiens1996). Although it is quite clear that night-time conductance (g~dark~) is much larger than a true 'minimum' conductance (discussed further in Section VI), it is frequently used as an independent measurement of g~0~ [@lombardozzi2017], and thus a comparison of magnitudes is useful. A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca. 150 species during night-time. Actual minimum daytime values were estimated from the @lin2015 database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset. This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted/extrapolated g~0~ term.

\
The results of the quantitative comparison are presented in Fig. 1. From this compilation, a few striking differences between estimates are apparent. Cuticular conductance averaged 0.45 mmol m^-2^ s^-1^, ten-fold lower compared to g~min~ (4.9 mmol m^-2^ s^-1^). This result is in line with the methods comparison of @kerstiens1996. More recently, @schuster2017 argued, based on a data for 8 species, that g~min~ from MLD is comparable to g~cut~. However in the same study, a literature compilation of hundreds of values showed that g~cut~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cut~ were pooled in their analysis. Nonetheless, the comparison between g~cut~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions. @boyer1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases.

\
Conductance in the dark (g~dark~) (mean = 41.4 mmol m^-2^ s^-1^ was on average eight-fold higher than g~min~. It is no new observation that stomata do not close in the dark (further discussed in Section VI). Previous studies have directly compared g~dark~ and g~min~ on the same species, and generally concluded that g~dark~ is much higher than g~min~ [@hygen1954; @koerner1994; @walden-coleman2013], but an exception is @cavender-bares2007 who found the two rates to be similar in oaks.

\
The fitted minimum conductance parameter, g~0~ (mean = 32.8 mmol m^-2^ s^-1^), was much higher than g~min~, but similar to g~dark~. However, we point out the uncertainties associated with g~0~ in Section III, which tend to inflate estimates of g~0~. It is telling that actual measured stomatal conductance during low photosynthesis (g~s,low A~) was  much lower than g~0~ or g~dark~. In other words, actual daytime conductance during these conditions that cause low photosynthetic rates is much lower than the presumed minimum estimated from fitted g~0~. A similar result was already presented by @dekauwe2015 (their Fig. A2).

@@

**Figure 1**. Comparison of various estimates of the (presumed) minimum conductance. Error bars are 95\% confidence intervals. Grey dots are the original data (but some data occurs outside the figure range). Different letters denote significant differences (at $\alpha$ = 0.05). g~cut,isol~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~0~ : fitted parameter (Eq. 1) with regression; g~s,low A~ leaf conductance during periods of very low photosynthesis. See text for data sources and methods.

@@

\
The standard approach in estimating g~0~ for use in Eq. 1 is to use regression with g~s~ as the response variable, and the right-hand side of Eq. 1 as the predictor. The g~0~ parameter is thus estimated as the intercept. Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated by this method. The difficulty of the estimation can already be anticipated from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g. @leuning1995 uses 0.01 mol m^-2^ s^-1^ for all species); presumably because unreliable estimates were obtained, b) negative g~0~ estimates from regression are commonly reported [@leuning1995; @heroult2013; @miner2017], though it is clearly non-sensical to suggest negative conductance values. @barnard2013 also mentioned the difficulty of fitting Eq. 1 to estimate g~0~, but did not present specific details - the following analysis builds on their work by demonstrating statistical uncertainties.

@@

\
In Fig. 2, we demonstrate how g~0~ and g~1~ are statistically correlated, when fitted for a single study. The correlation between slopes and intercepts is not unique to Eq. 1, but a general property of linear models [@stapleton1995]. The consequence is that it is more difficult to interpret g~1~ and g~0~ independently, and that confidence intervals on each of the parameters should also not be interpreted independently (see @becker2007 for more discussion on the statistical aspects of comparing slopes and intercepts).

@@

**Figure 3**. Statistical uncertainty in the estimation of g~0~ from regression, demonstrated with two parameter databases. (a) We fitted the linearized form of the @medlyn2011 model to each of the datasets in the @lin2015 leaf gas exchange database, showing that for poorly fitted relationships (low R^2^), inflated estimates of g~0~ are obtained.  Vertical lines are 95% confidence intervals. The grey line is a fitted loess smoother with 95% confidence interval. Note the wide confidence intervals, and frequent negative values. (b) Similar to (a), but using the published compilation by @miner2017. The grey line is a fitted loess smoother with 95% confidence interval. (c) Using the fits from panel (a), a demonstration that the standard error (SE) of g~0~ is much higher when the coefficient of variation (CV) of the predictor (i.e. right-hand side of the equation being fitted) is lower.

@@

\
In this review we focus on the cuticle as a barrier to water loss, but point out that previous work has shown that the cuticle is much less permeable to CO~2~ compared to H~2~O [@boyer1997; @boyer2015]. This finding, if indeed generally true, has a large effect on the calculation of various gas exchange parameters [@hanson2006]. @manzoni2011 recalculated leaf water-use efficiency for droughted plants assuming CO~2~ is blocked by the cuticle, but the uncertainty of this approach is leaky and incompletely closed stomata. Because g~0~ includes not just the cuticle, but also incompletely closed stomata, and the finding that g~0~ is much larger than estimates of g~min~ or g~cut~ (Fig. 1), we ignore this effect in the model simulations that follow.

\
Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero. But A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, begging the question whether the same g~0~ is reached in each of these cases. In the application of the Ball-Berry models, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005]. It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cut~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata. Photosynthesis also ceases at very high temperature, but stomata do not appear to always close in proportion to this decrease in photosynthesis, if at all [@drake, @urban2017].

@@

\
Using the coupled leaf gas exchange model, we show in Fig. 4 how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^). We also demonstrate the effect of g~0~ on modelling the T~leaf~ response. In this simulation, T~leaf~ and VPD are assumed to co-vary with an empirical relationship as used by @duursma2014. Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increase leaf transpiration E~L~, but at a slightly higher rate than just due to g~0~. This effect arises because the additional conductance allows slightly higher rates of photosynthesis, which in turn increase g~s~ via Eq. 1.

@@

**Figure 4**. Simulations with a coupled leaf gas exchange model [@duursma2015], demonstrating the effect of inclusion of the g~0~ parameter (Eq. 1) on leaf fluxes. (a) Intrinsic water-use efficiency (A~n~/g~s~) as a function of he photosynthetic photon flux density (PPFD), holding other environmental drivers constant, for three values of g~0~. (b) The same simulations as in (a), but showing the intercellular CO~2~ concentration (C~i~). (c) Stomatal conductance (g~s~) simulations, where vapour pressure deficit (VPD) and air temperature (T~air~) were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions. (d) The same simulations as in (c), but showing C~i~. Note how C~i~ increases at high VPD and T~air~, only when g~0~> 0. For all simulations it is assumed that T~leaf~ is equal to T~air~, and ignoring the differential permeability of the cuticle to CO~2~ and H~2~O [@hanson2016].

## 2. Models of plant desiccation

\
When plants are sufficiently water-stressed that stomata are mostly closed, water loss still continues at a rate determined by the minimum conductance. Thus models that aim to predict when plants desiccate and die must include a minimum conductance term. A classic study by @pisekwinkler1953 already calculated the length of time needed to desiccate leaves to some critical low water content, given the minimum transpiration rate, and the saturated water content of the leaves. Based on that work, @burghardt2006 reported on a direct correlation between g~min~ and the survival time of leaves. @sinclair2000 presented the minimum conductance as a key drought tolerance trait, and used it to as a basis for predicting crop mortality during severe drought. More recently, @gleason2014 and @blackman2016 proposed that embolism resistance along with whole plant capacitance and minimum transpiration rates all contribute to define the time to desiccation. Finally, @martinstpaul2017 demonstrated in a whole plant model of hydraulic failure that g~min~ was one of the key parameters to explain the drop in water potential below the cavitation threshold, because stomata generally close well before this threshold. Relying on the *Sureau* model presented in @martinstpaul2017 we illustrate the critical role of g~min~ in defining the desiccation time (Fig. 5).

@@

**Figure 5**. Simulations with the *Sureau* model demonstrating the effect of g~min~ on desiccation tolerance of plants. The *Sureau* model simulates water transport in the soil-plant-atmosphere continuum, and includes a detailed representation of capacitance in stem and leaf tissues. (a) Soil relative extractable water (REW; 1 = field capacity, 0 = permanent wilting point) for the two simulations, using a minimum conductance (g~min~) of 1.6 or 3.3 mmol m^-2^ s^-1^ - all other parameters were equal. (b) Water potential in the soil and leaf as the dry-down progresses. Note the brief increase in soil water potential around days 40 - 70; this represents water released from stem cavitation. (c) Progression of percent loss conductivity (PLC) of the xylem. If we assume that plant mortality occurs at a PLC of 88%, the plants with the low g~min~ lasted 15 days longer (or nearly 40%).

@@

\
A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, as these processes have been well described elsewhere [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b]. However, a few key points should be summarized as they are relevant to the current discussion, in particular when we aim to interpret variation in literature values of g~min~ (Section VII).

\
Although we use the term cuticular conductance freely, this transport pathway does not represent a true conductance since water does not diffuse as a gas through the cuticle. Instead, it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle [@kerstiens1996; @riederer2001; @schreiber1996]. The main barrier to diffusion is actually a very thin layer at the leaf surface. Because most of the resistance is located in such a thin layer, g~cut~ does not correlate with the thickness of the cuticle [@priestley1943; @riederer2001; @anfodillo2002; @schuster2016b]. Nonetheless there is considerable variation in cuticle thickness between plant species [@schuster2016a], as well as within species in response to drought (**cite**), along altitudinal transects [@delucia1984], and even with increasing height in the canopy of very tall trees [@woodruff2010]. If the cuticle thickness does not directly affect its conductance, what is the value of increasing the cuticle?

\
It is important to bear in mind the many other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, various interactions with insects including signalling and herbivory resistance [@mueller2008; @kerstiens1996c;@riederer2008], and foliar water uptake [@fernandez2017] The thickness of the cuticle may also confer mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness). As such, thick cuticles may be advantageous because prolonged exposure to wind increases the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991], by dislodging cuticular compounds by abrasion [@rogge1993]. Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986], most dramatically demonstrated in an ice storm, leading to much higher g~min~ [@boyce2003]. Such exposure to the elements was also held responsible for thicker cuticles in coastal, wind-exposed sites compared to inland sites [@blackman2005].
